Reparametrization of COM-Poisson Regression Models with Applications in the Analysis of Experimental Data

### 2018-01-31 article.Rnw
Do minor corrections and add submission version of the paper

% -------------------------------------------
% Authors and affiliations
\Author{Eduardo E. Ribeiro Jr\Affil{1},
  Walmes M. Zeviani\Affil{2},
  Wagner H. Bonat\Affil{2},
  Clarice G. B. Dem\'{e}trio\Affil{1}
  and John Hinde\Affil{3}
}
\AuthorRunning{Ribeiro Jr \textrm{et al.}}

@@

% -------------------------------------------
% Abstract
\Abstract{ In the analysis of count data often the equidispersion
  assumption is not suitable, hence the Poisson regression model is
  inappropriate.  As a generalization of the Poisson distribution the
  COM-Poisson distribution can deal with under-, equi- and overdispersed
  count data.  It is a member of the exponential family of distributions
  and has the Poisson and geometric distributions as special cases, as
  well as the Bernoulli distribution as a limiting case.  In spite of
  the nice properties of the COM-Poisson distribution, its location
  parameter does not correspond to the expectation, which complicates
  the interpretation of regression models specified using this
  distribution. In this paper, we propose a straightforward
  reparametrization of the COM-Poisson distribution based on an
  approximation to the expectation of this distribution. The main
  advantage of our new parametrization is the straightforward
  interpretation of the regression coefficients in terms of the
  expectation of the count response variable, as usual in the context of
  generalized linear models. Furthermore, the estimation and inference
  for the new COM-Poisson regression model can be done based on the
  likelihood paradigm. We carried out simulation studies to verify the
  finite sample properties of the maximum likelihood estimators. The
  results from our simulation study show that the maximum likeli-hood
  estimators are unbiased and consistent for both regression and
  dispersion parameters. We observed that the empirical correlation
  between the regression and dispersion parameter estimators is close to
  zero, which suggests that these parameters are orthogonal. We
  illustrate the application of the proposed model through the analysis
  of three data sets with over-, under- and equidispersed count data.
  The study of distribution properties through a consideration of
  dispersion, zero-inflated and heavy tail indices, together with the
  results of data analysis show the flexibility over standard
  approaches. Therefore, we encourage the application of the new
  parametrization for the analysis of count data in the context of
  COM-Poisson regression models.  The computational routines for fitting
  the original and new version of the COM-Poisson regression model and
  the analyzed data sets are available in the supplementary material.
}

@@

To illustrate the orthogonality, \autoref{fig:ortho-surf} displays
contour plots of the deviance surfaces for four simulated data set of
size 1000, $\mu=5$ and different values of the dispersion
parameters. The shapes of the deviance function show that the proposed
parametrization is better for both computation and asymptotic
(normal-based) inference. Furthermore, it is interesting to note that
the deviance function shape under strong overdispersion ($\phi=-1.6$) is
not as well behaved as the others; this is due to the difficulty of the
distribution in dealing with strong overdispersion in low counts (see
dispersion index plot in the \autoref{fig:indexes-plot}). This also
explains the results of Cov$(\hat{\beta}_{22}, \phi)$ in the first panel
of \autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and
associated with low counts.

>>>>

% -------------------------------------------
% Authors and affiliations
\Author{Eduardo E. Ribeiro Jr\Affil{1},
  Walmes M. Zeviani\Affil{2},
  Wagner H. Bonat\Affil{2},
  Clarice G. B. Dem\'{e}trio\Affil{1}
  and John Hinde\Affil{3}
}
\AuthorRunning{Eduardo E. Ribeiro Jr \textrm{et al.}}

@@

% -------------------------------------------
% Abstract
\Abstract{ In the analysis of count data often the equidispersion
  assumption is not suitable, hence the Poisson regression model is
  inappropriate.  As a generalization of the Poisson distribution the
  COM-Poisson distribution can deal with under-, equi- and overdispersed
  count data.  It is a member of the exponential family of distributions
  and has the Poisson and geometric distributions as special cases, as
  well as the Bernoulli distribution as a limiting case.  In spite of
  the nice properties of the COM-Poisson distribution, its location
  parameter does not correspond to the expectation, which complicates
  the interpretation of regression models specified using this
  distribution. In this paper, we propose a straightforward
  reparametrization of the COM-Poisson distribution based on an
  approximation to the expectation of this distribution. The main
  advantage of our new parametrization is the straightforward
  interpretation of the regression coefficients in terms of the
  expectation of the count response variable, as usual in the context of
  generalized linear models. Furthermore, the estimation and inference
  for the new COM-Poisson regression model can be done based on the
  likelihood paradigm. We carried out simulation studies to verify the
  finite sample properties of the maximum likelihood estimators. The
  results from our simulation study show that the maximum likeli-hood
  estimators are unbiased and consistent for both regression and
  dispersion parameters. We observed that the empirical correlation
  between the regression and dispersion parameter estimators is close to
  zero, which suggests that these parameters are orthogonal. We
  illustrate the application of the proposed model through the analysis
  of three data sets with over-, under- and equidispersed count data.
  The study of distribution properties through a consideration of
  dispersion, zero-inflated and heavy tail indexes, together with the
  results of data analysis show the flexibility over standard
  approaches. Therefore, we encourage the application of the new
  parametrization for the analysis of count data in the context of
  COM-Poisson regression models.  The computational routines for fitting
  the original and new version of the COM-Poisson regression model and
  the analyzed data sets are available in the supplementary material.
}

@@

To illustrate the orthogonality, \autoref{fig:ortho-surf} displays
contour plots of the deviance surfaces for four simulated data set of
size 1000 with $\mu=5$ and different values of the dispersion
parameters. The shapes of the deviance function show that the proposed
parametrization is better for both computation and asymptotic
(normal-based) inference. Furthermore, it is interesting to note that
the deviance function shape under strong overdispersion ($\phi=-1.6$) is
not as well behaved as the others; this is due to the difficulty of the
distribution in dealing with strong overdispersion in low counts (see
dispersion index plot in the \autoref{fig:indexes-plot}). This also
explains the results of Cov$(\hat{\beta}_{22}, \phi)$ in the first panel
of \autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and
associated with low counts.
